Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-10 manuscript.tex
starting delay on chip
Update on Overleaf.

\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e. the amount of myelin wrapped around the axon. Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez_cerda_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}. Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def). Each sheath \citep{simons_2016}.

Thus in several occasions, myelin thus has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. delay regulation.

@@

facilitates both the neural circuit function and the behavioural performance~\cite{fields_2015}. Experiments on mammals show that myelination is directly related to learning and memory consolidation, especially motor training and in enriched environments. This biological phenomenom both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations. It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.

@@

Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.

@@

Artificial intelligence has made huge advances in the past decades, now deep learning algorithms outperform humans at complex tasks like natural image recognition or abstract strategy board games. Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency. Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy2018toward}). Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations. This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance. These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation for the recording of a visual scene (ref). Another important distinction with classical Von Neumann architectures is the localized memory of this new type of chips. It is materialized by capacity of the physical connections between the processing units to store information. An example of such connection, directly inspired by synaptic plasticity, is the memristor (ref). Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems. Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (zenke2021visualizing). Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER). We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.

@@

>>>>

\citet{gasser_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e. the amount of myelin wrapped around the axon. Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez_cerda_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmidt_1939}. Myelination consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\cite{simons_2016}. In average, each oligodendrocyte produce 20 to 60 myelinating processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_2016}. Multiple sheaths of myelin can appear on one neuron axon and form the internode between the Ranvier nodes (def); this demonstrate an additional impact of myelination on the conduction velocity as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}. Thus in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. axonal delay regulation.

@@

\citet{fields_2015} states that myeline facilitates both the neural circuit function and the behavioural performance: experiments on mammals show that myelination is activity-dependant and directly related to learning and memory consolidation, especially motor training and in enriched environments. This biological phenomenom takes place both at an early age, where the amount of oligodendrocites is the particularly high in the central nervous system~\citep{reynolds_1928}, and in older animals, due to its involvement in coupling the activity of distant neuron populations. It is worth highlighting that myelination becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions.

@@

However, one question remains: how do oligodendrocytes detect neuronal activity and regulate the myelination accordingly? To answer this, we must first study the myelination process. The oligodendrocyte precursor cells (OPCs) first proliferate in the white matter \textit{via} a self-repursive process, thus allowing for an evenly spaced network, and identify target axons. Most OPCs then differentiate into oligodendrocytes and immediatly initiate myelination, with no further migration.

axo-glial signaling + how do they identify the target axon

Extensive studies of the glialo axo-glial synaptic communication
Some further studies show that oligodendrocytes may detect neuronal activity thanks to growth factors or neurotransmitters released through ion channels or via exocytosis.

% myelination for axonal delay, but what about dendritic delay ? 
The previous paragraphs present the biological mechanisms behind axonal delay. However dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes. cite all info from \citep{MaladiAsl_2018} + \citep{branco_dendritic_2010}

@@

Artificial intelligence has made huge advances in the past decades, now deep learning algorithms outperform humans at complex tasks like natural image recognition or abstract strategy board games. Machine learning algorithms still suffer from adversarial attacks or a lack of generalization capacity but their main weakness, compared to biological neural networks, rely on their poor energy efficiency. Neuromorphic engineering intends to mimic the neural bases of communication with a wide variety of technics, from strictly analog circuits to software-based neuromorphic systems, and to develop tools improving the capacities of current artificial intelligence (\cite{roy2018toward}). Because the reduced energy consumption of biological networks can be explained by the use of spikes and asynchronous responses to exchange information (ref), neuromorphic chips use this parallel and event-based representation to make energy efficient computations. This is also the case for Dynamic Vision Sensors (DVS) which provide a stream of binary asynchronous events signaling detectable changes in luminance. These devices, also named ``silicon retinas'' (see Figure \ref{fig:silicon_retina}), show great improvements in terms of memory allocation for the recording of a visual scene (ref). Another important distinction with classical Von Neumann architectures is the localized memory of this new type of chips. It is materialized by capacity of the physical connections between the processing units to store information. An example of such connection, directly inspired by synaptic plasticity, is the memristor~\citep{markovic_physics_2020}. Using these event-based computations as a building brick, neuromorphic engineering proposes new hardware designs perfect to simulate SNNs and use the full power of asynchronous computations observed in biological systems. Many connections can be drawn between neuromorphic engineering and computational neuroscience to aim at solving both research and technology challenges (zenke2021visualizing). Just as communication between neurons can be illustrated by a raster plot, information exchanges in neuromorphic hardware are described as event streams by an Address-Event Representation (AER). We will use indifferently \textit{events} or \textit{spikes} to describe the same binary and discrete processes and, in this section, we will extend the review of the use of precise spike timing to this emerging field.

@@

\citep{renner_sparse_2022}

\citep{sandamirskaya_neuromorphic_2022}

\citep{hussain_deltron_2012}
\citep{sandamirskaya_neuromorphic_2022}
\citep{sandamirskaya_neuromorphic_2022}

%https://ieeexplore.ieee.org/abstract/document/8123631

%https://ieeexplore.ieee.org/abstract/document/9622153

%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160

%Sparse Vector Binding on Spiking Neuromorphic Hardware Using Synaptic Delays
